---
title:"Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
=====================================================


## Loading and preprocessing the data
##Assuming that the file is already unzipped in the local folder
```{r,echo=TRUE}
data <- read.csv("activity.csv",header=T)
```


## What is mean total number of steps taken per day?
```{r,echo=TRUE}
data_trim <- na.omit(data) # Considering only complete cases
steppday <- aggregate(steps ~ date, data_trim, sum) #calculating total steps per day
hist(steppday$steps,col=1,main="Histogram - Total No. of Steps per Day",xlab="Total No. of steps each day")
```
## Mean & Median of Total No. of Steps taken per Day
```{r,echo=TRUE}
sprintf("Mean Value for Total No. of Steps per Day is '%a'",mean(steppday$steps))
sprintf("Median Value for Total No. of Steps per Day is '%a'",median(steppday$steps))
```


## What is the average daily activity pattern?
```{r,echo=TRUE}
int_steps <- aggregate(steps ~ interval, data_trim, mean) #Calculating Steps with respect to interval and summing them up
plot(int_steps$interval,int_steps$steps,type='l',col=1,main="Avg. No. of Steps across all Days",xlab="Interval",ylab="Avg. No. of Steps")
```
## Interval with Max. No. of Steps across all the Days
```{r,echo=TRUE}
max_steps_int <- int_steps[which.max(int_steps$steps),] ##Finding row with max interval
sprintf("Max interval is '%a'-'%a'with max. avg. no. of steps'%a'",as.numeric(max_steps_int$interval),as.numeric(max_steps_int$interval)+5,as.numeric(max_steps_int$steps))
```


## Imputing missing values
```{r,echo=TRUE}
data_na <- data[!complete.cases(data),] ##Rows with NAs
sprintf("No.of rows with missing values are '%a'",nrow(data_na))
```
##Impute Mean for NA Values of the 5-min interval
```{r,echo=TRUE}
for(i in 1:nrow(data)){
        if(is.na(data$steps[i])){
                int_val <- data$interval[i]
                rid <- which(int_steps$interval==int_val)
                steps_val <- int_steps$steps[rid]
                data$steps[i] <- steps_val
        }
} 
```
##Aggregate Steps with respect to date
```{r,echo=TRUE}
imputed_steps <- aggregate(steps ~ date,data,sum)
```
##Creating Histogram with Imputed Values
```{r,echo=TRUE}
hist(imputed_steps$steps,col=1,main="Histogram with Imputed Values",xlab="Total No. of Steps each day")
```
##Calcuating Mean & Median for Original Data Set and Data Set with Imputed Values
```{r,echo=TRUE}
sprintf("Original Data Set Mean Value - '%a'",mean(steppday$steps))
sprintf("Original Data Set Median Value - '%a'",median(steppday$steps))
sprintf("Imputed Data Set Mean Value - '%a'",mean(imputed_steps$steps))
sprintf("Imputed Data Set Median Value - '%a'",median(imputed_steps$steps))
```
* Though the Means remain the same due to imputation, there is minor difference (by 1) in Median


## Are there differences in activity patterns between weekdays and weekends?
```{r,echo=TRUE}
#Convert Date to Date Class
data$date <- as.Date(data$date,"%Y-%m-%d")
data$day <- weekdays(data$date) # Including Day of the Week
data$day_type <- c("weekday") # Including new varible Daytype
for(i in 1:nrow(data)) { # Segregating Weekday & Weekends
        if(tolower(data$day[i]) == "saturday" || tolower(data$day[i]) == "sunday") {
                data$day_type <- c("weekend")
        }
}
#Convert Day Type to Factor Variable
data$day_type = as.factor(data$day_type)
# Aggregate Steps across interval and Day Type
imp_int_steps <- aggregate(steps ~ interval + day_type, data, mean)
#Call ggplot2 package for panel plots
library(ggplot2)
qplot(interval, steps, data=imp_int_steps, geom=c("line"),xlab="Interval",ylab="No. of Steps",main="")+facet_wrap(~day_type,ncol=1)
rm(list=ls()) # Frees the memory
#library(knitr)
#knit2html("PA1_template.Rmd",keep_md=TRUE)
```